\begin{acexercise}\end{acexercise}

Consider a multiprocessor with a symmetric shared memory architecture 
with snooping protocol. 
Each processor has a cache where consistency is kept by using the MSI protocol.
Each cache block has a single word.

The following table shows the initial state of four different variables in
each of the caches.

\medskip

\begin{tabular}{|c|c|c|c|c|}
\hline
&
\multicolumn{4}{|c|}{Initial states of the variables}
\\
\hline

\textbf{Processor} &
\textbf{A} &
\textbf{B} &
\textbf{C} &
\textbf{D}
\\
\hline
\hline

\textbf{P0} & 
Shared & Modified & Shared & Shared
\\
\hline

\textbf{P1} &
Invalid & Invalid & Invalid & Shared
\\
\hline

\textbf{P2} &
Invalid & Invalid & Shared & Shared
\\
\hline

\end{tabular}

\medskip

The following table shows the final state of these variables after doing a
series of accesses to memory.

\medskip

\begin{tabular}{|c|c|c|c|c|}
\hline
&
\multicolumn{4}{|c|}{Final states of the variables}
\\
\hline

\textbf{Processor} &
\textbf{A} &
\textbf{B} &
\textbf{C} &
\textbf{D}
\\
\hline
\hline

\textbf{P0} & 
Invalid & Invalid & Invalid & Shared
\\
\hline

\textbf{P1} &
Invalid & Invalid & Shared & Modified
\\
\hline

\textbf{P2} &
Modified & Modified & Invalid & Shared
\\
\hline

\end{tabular}

\medskip

Complete the following:

\begin{itemize}

  \item For each variable (A, B, C and D) describe the
accesses made and the processes involved that have allowed to reach
the final state.
 

    \begin{itemize}

      \item \textbf{Note 1}: to reach the final state may be enough
a single access or a sequence of accesses.

      \item \textbf{Note 2}: there may be a final state
that is unreachable (that is, that has no solution).

    \end{itemize}

  \item For each previous case describe the generated bus traffic.

\end{itemize}

\begin{acsolution}\end{acsolution}

\begin{itemize}

\item Variable \cppid{A}:
  \begin{itemize}
    \item \textbf{\textbf{P2}}  writes \cppid{A} (\emph{exclusive}) invalidating \textbf{P0}.
    \item \textbf{P2} produces a write miss that is captured by \textbf{P0} and produces no traffic.
  \end{itemize}

\item Variable \cppid{B}:
  \begin{itemize}
    \item \textbf{P2}  writes \cppid{B}  (\emph{exclusive}) invalidating the copy of \textbf{P0}
    \item \textbf{P2} produces a write miss that is captured by \textbf{P0} that performs a write-block to the cache of \textbf{P2}
  \end{itemize}

\item Variable \cppid{C}:
  \begin{itemize}
    \item \textbf{P1} writes \cppid{B} (\emph{exclusive}) invalidating the other two copies
    \item \textbf{P1} produces a write miss that is captured by \textbf{P0} and \textbf{P1}. No bus traffic is generated.
    \item \textbf{P1} reads of writes in \cppid{D} another variable with a related block allocated in the same line as \cppid{C} producing a conflict miss. This makes block \cppid{C} to be replaced and given it was modified it is written into memory. Block \cppid{C} is not present in cache (equivalent to \emph{invalid} state).
    \item This action of \textbf{P1} produces a \emph{write-back} of the block in memory. 
    \item \textbf{P1} reads \cppid{C}. Produces a read miss and goes to shared state. 
  \end{itemize}

\item Variable \cppid{D}:
  \begin{itemize}
    \item The final state can not be reached because a block can not be exclusive and shared at the same time.
  \end{itemize}

\end{itemize}
