\section{Introduction}

\begin{frame}[t]{Latency evolution}
\begin{itemize}
  \item Multiple views of performance
    \begin{itemize}
      \item $performance = \frac{1}{latency}$
      \item Useful for comparing processor and memory evolution.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Processors}
    \begin{itemize}
      \item Yearly performance increase from 25\% to 52\%.
      \item Combined effect from 1980 to 2005 $\rightarrow$ above $10,000$ times.
      \item Marginal growth since then.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Memory}
    \begin{itemize}
      \item Yearly performance increase around 7\%
      \item Combined effect from 1980 to 2015 $\rightarrow$ around $10$ times.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Multi-core effect}
\begin{itemize}
  \item \textmark{Intel Core i7 6700}.
    \begin{itemize}
      \item Two 64 bits data accesses per cycle.
      \item 4 cores, 4.2 GHz $\rightarrow$ $32.8 \times 10^9$ references/sec
      \item Instructions demand: $12.8 \times 10^9$ of 128 bits.
      \item Peak bandwidth: $524.8$ GB/sec + $204.8$ GB/s
    \end{itemize}

  \mode<presentation>{\vfill}
  \item \textmark{SDRAM memory}.
    \begin{itemize}
      \item DDR2 (2003): 3.20 GB/sec – 8.50 GB/sec
      \item DDR3 (2007):  8.53 GB/sec – 18.00 GB/sec
      \item DDR4 (2014): 17.06 GB/sec – 25.60 GB/sec
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Solutions}:
    \begin{itemize}
      \item Multi-port memories, pipelined caches, multi-level caches,
            per-core caches, instruction/data caches separation.
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t]{Principle of locality}
\begin{itemize}
  \item \textgood{Principle of locality}.
    \begin{itemize}
      \item It is \textmark{property of programs} exploited in the hardware design.
      \item Programs are accessed in a relatively small portion of address space.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Types of locality}:
    \begin{itemize}
      \item \textmark{Temporal locality}: 
            Elements recently accessed tend to be accessed again.
        \begin{itemize}
          \item Examples: Loops, variable reuse, \ldots
        \end{itemize}
      \item \textmark{Spatial locality}: 
            Elements next to a recently accessed one tend to be accessed in the future.
        \begin{itemize}
          \item Examples: Sequential execution of instructions, arrays, \ldots
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Memory hierarchy}
\makebox[\textwidth][c]{
\input{en/m3-01-cache/mem-hier.tkz}
}
\end{frame}

\begin{frame}[t]{Situation (2017)}
\begin{itemize}
  \item \textgood{Registers} $\rightarrow$ \emph{CMOS}.
    \begin{itemize}
      \item \textmark{Access Time}: 0.1 ns -- 0.2 ns.
      \item \textmark{Typical size}: < 4KiB.
      \item \textmark{Bandwidth}: $10^6$ -- $10^7$ MiB/s
    \end{itemize}

  \item \textgood{Cache} $\rightarrow$ \emph{On-chip CMOS SRAM}.
    \begin{itemize}
      \item \textmark{Access time}: 0.5 ns -- 1.0 ns.
      \item \textmark{Typical size}: 32KiB -- 8 MiB.
      \item \textmark{Bandwidth}: $2 \cdot 10^4$ -- $5 \cdot 10^4$ MiB/s 
    \end{itemize}

  \mode<presentation>{\vfill}
  \item \textgood{Main memory} $\rightarrow$ \emph{CMOS DRAM}.
    \begin{itemize}
      \item \textmark{Access time}: $30$ ns -- $150$ ns
      \item \textmark{Typical size}: < 1 TB
      \item \textmark{Bandwidth}: $10^4$ -- $3 \cdot 10^4$ MiB/s 
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t]{Scale of latencies}
\begin{itemize}
  \item \textgood{Tiempos típicos}:
    \begin{itemize}
      \item \textmark{CPU Cycle}: 0.3 ns
      \item \textmark{L1}: 0.9 ns 
      \item \textmark{L2}: 2.8 ns 
      \item \textmark{Main memory}: 120 ns 
      \item \textmark{SSD}: 50--150 $\mu$s 
      \item \textmark{Disco}: 1--10 ms
    \end{itemize}

  \pause
  \item \textgood{Si un ciclo fuese un segundo}:
    \begin{itemize}
      \item \textmark{Ciclo de CPU}: 1 s
      \item \textmark{Caché L1}: 3 s
      \item \textmark{Caché L2}: 9 s
      \item \textmark{Memoria principal}: 6 min.
      \item \textmark{SSD}: 2--6 días.
      \item \textmark{Disco}: 1-12 meses.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Memory hierarchy}
\begin{itemize}
  \item \textgood{Block or line}: Unit of copy operations.
    \begin{itemize}
      \item Usually composed of multiple words.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item If accessed data is \textmark{present} in upper level:
    \begin{itemize}
      \item \textgood{Hit}: Delivered by higher level.
    \end{itemize}
\[
h = \frac{hits}{acceses}
\]

  \mode<presentation>{\vfill\pause}
  \item If accessed data is \textmark{missing}.
    \begin{itemize}
      \item \textgood{Miss}: Block copied from lower level.
      \item Data access in upper level.
      \item Needed time $\rightarrow$ \textmark{Miss penalty}.
    \end{itemize}
\[
m = \frac{misses}{acceses} = 1 - h
\]

\end{itemize}
\end{frame}

\begin{frame}[t]{Metrics}
\begin{itemize}
  \item Average memory access time:
\[
t_M = t_H + (1-h) t_M
\]
  \mode<presentation>{\vfill\pause}
  \item \textgood{Miss penalty}:
    \begin{itemize}
      \item Time to replace a block and deliver to CPU.
      \item \textmark{Access time}.
        \begin{itemize}
          \item Time to get from lower level.
          \item Depends on lower level latency.
        \end{itemize}
      \item \textmark{Transfer time}.
        \begin{itemize}
          \item Time to transfer a block.
          \item Depends on the bandwidth across levels.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Metrics}
\begin{itemize}
  \item \textgood{CPU execution time}:
\[
t_{CPU} = 
\left( cycles_{CPU} + cycles_{\text{memory stall}} \right) \times t_{cycle}
\]

  \mode<presentation>{\vfill\pause}
  \item \textgood{CPU clock cycles}:
\[
cycles_{CPU} =
IC \times CPI
\]

  \mode<presentation>{\vfill\pause}
  \item \textgood{Memory stall cycles}:
\[
cycles_{\text{memory stall}} =
n_{misses} \times \text{penalty}_{miss}=
\]
\[
IC \times miss_{instr} \times \text{penalty}_{miss} =
\]
\[
IC \times memory\_accesses_{instr} \times (1 - h ) \times \text{penalty}_{miss}
\]
\end{itemize}
\end{frame}

\begin{frame}[t]{Difference between reads and writes}
\begin{itemize}
  \item Read and write different patterns:
    \begin{itemize}
      \item Differen hit rates: $h_{read}$, $h_{write}$.
      \item Different penalties: $\text{penalty}_{read}$, $\text{penalty}_{write}$
    \end{itemize}

\[
ciclos_{\text{detención memoria}} =
IC \times reads_{instr} \times (1 - h_{read}) \times \text{penalty}_{read} +
\]
\[
IC \times writes_{instr} \times (1 - h_{write}) \times \text{penalty}_{write}
\]
\end{itemize}
\end{frame}

