\section{Introduction}

\begin{frame}[t]{What is OpenMP?}
\begin{itemize}
  \item It is an \textmark{language extension} for expressing parallel applications
        in shared memory systems.

  \mode<presentation>{\vfill\pause}
  \item \textgood{Components}:
    \begin{itemize}
      \item Compiler directives.
      \item Library functions.
      \item Environment variables.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Simplifies the way of writing parallel programs.
    \begin{itemize}
      \item \emph{Mappings} for FORTRAN, C and C++.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Constructs}
\begin{itemize}
  \item Directives:
\begin{lstlisting}
#pragma omp directive [clause]
\end{lstlisting}

  \begin{itemize}
    \item \textmark{Example}: Setup the number of threads.
\begin{lstlisting}
#pragma omp parallel num_threads(4)
\end{lstlisting}
  \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Library functions:
\begin{lstlisting}
#include <omp.h> // Include to call OpenMP  API functions
\end{lstlisting}
    \begin{itemize}
      \item \textmark{Example}: Get the number of threads in use.
\begin{lstlisting}
int n = omp_get_num_threads();
\end{lstlisting}
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Exercise 1: Sequential}
\begin{block}{ex1seq.cpp}
\lstinputlisting{lab/04-omp/hello/exseq.cpp}
\end{block}
\begin{itemize}
  \item Print to standard output.
\end{itemize}

\begin{lstlisting}[style=terminal]
Hello(0) world(0)
\end{lstlisting}
\end{frame}

\begin{frame}[t,fragile]{Exercise 1: Parallel}
\begin{columns}[T]

\column{.5\textwidth}
\begin{block}{ex1par.cpp}
\lstinputlisting{lab/04-omp/hello/expar.cpp}
\end{block}

\column{.5\textwidth}

\begin{itemize}
  \item Print to standard output.
\end{itemize}

\begin{lstlisting}[style=terminal]
Hello(Hello(10) world(0Hello() world(1)
3) world(3)
Hello(2) world(2)
)
\end{lstlisting}

\end{columns}
\end{frame}

\begin{frame}[t,fragile]{Compiling OpenMP}

\begin{itemize}
  \item Compiler flags: 
    \begin{itemize}
      \item \textgood{gcc}: \textmark{-fopenmp}
      \item \textgood{Intel Linux}: \textmark{-openmp}
      \item \textgood{Intel Windows}: \textmark{/Qopenmp}
      \item \textgood{Microsft Visual Studio}: \textmark{/openmp}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Usando \textemph{CMake}:
  \begin{itemize}
    \item Activar el soporte para \textmark{OpenMP} en CMake.
\begin{lstlisting}
find_package(OpenMP REQUIRED)
\end{lstlisting}

    \item Activar las opciones de compilación y enlace para un ejecutable.
\begin{lstlisting}
add_executable(expar expar.cpp)
target_link_libraries(expar PUBLIC OpenMP::OpenMP_CXX)
\end{lstlisting}

  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Exercise 1}
\begin{itemize}
  \item \textgood{Goal}: Verify you have a working environment.
  
  \mode<presentation>{\vfill\pause}
  \item \textgood{Activities}:
  \begin{enumerate}
    \item Compile sequential version and run.
    \item Compile parallel version and run.
    \item Add a call to function \cppid{omp\_get\_num\_threads()} 
          to print the number of threads:
      \begin{enumerate}[a)]
        \item Before the \cppkey{pragma}.
        \item Just after \cppkey{pragma}.
        \item Within the block.
        \item Before exiting the program, but outside the block.
      \end{enumerate}
  \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}[t]{Observations}
\begin{itemize}
  \item A model for multi-threaded shared memory.
    \begin{itemize}
      \item Communication through shared variables.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Accidental sharing $\rightarrow$ \textbad{race conditions}.
    \begin{itemize}
      \item Result depending on threads scheduling.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Avoiding race conditions.
    \begin{itemize}
      \item Synchronize to avoid conflicts.
        \begin{itemize}
          \item Cost of synchronizations.
        \end{itemize}
      \item Modify access pattern.
        \begin{itemize}
          \item Minimize needed synchronizations.
        \end{itemize}
    \end{itemize}

\end{itemize}
\end{frame}
