\section{Synchronization}

\begin{frame}[t]{Synchronization mechanisms}
\begin{itemize}
  \item \textgood{Synchronization}: 
        Mechanism used to establish constraints on the access order
        to shared variables.
    \begin{itemize}
      \item \textmark{Goal}: Avoid data races.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Alternatives}:
    \begin{itemize}
      \item \textgood{High level}: \cppkey{critical}, \cppkey{atomic}, \cppkey{barrier}, \cppkey{ordered}.
      \item \textgood{Low level}: \cppkey{flush}, lock.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{\emph{critical}}
\begin{itemize}
  \item Guarantees \textmark{mutual exclusion}.
  \item Only a thread at a time can enter the critical region.
\end{itemize}

\mode<presentation>{\vfill\pause}
\begin{columns}

\column{.35\textwidth}
\begin{block}{Example}
\begin{lstlisting}
#pragma omp parallel
{
  for (int i=0;i<max;++i) {
    x = f(i);
    #pragma omp critical
    g(x);
}
\end{lstlisting}
\end{block}

\column{.65\textwidth}
\begin{itemize}
  \item Calls to \cppid{f()} are performed in parallel.
  \item Only a thread can enter function \cppid{g()} at a time.
\end{itemize}

\end{columns}
\end{frame}

\begin{frame}[t,fragile]{\emph{atomic}}
\begin{itemize}
  \item Guarantees \textmark{atomic update} of a single memory location.
  \item Avoid data races in variable update.
\end{itemize}

\mode<presentation>{\vfill\pause}
\begin{columns}

\column{.35\textwidth}
\begin{block}{Example}
\begin{lstlisting}
#pragma omp parallel
{
  for (int i=0;i<max;++i) {
    x = f(i);
    #pragma omp atomic
    s += g(x)
}
\end{lstlisting}
\end{block}

\column{.65\textwidth}
\begin{itemize}
  \item Calls to \cppid{f()} performed in parallel.
  \item Updates to \cppid{s} are \emph{thread-safe}.
\end{itemize}

\end{columns}
\end{frame}

\begin{frame}[t]{Exercise 3}
\begin{itemize}
  \item Modify program from exercise 2.
  \item Evaluate:
    \begin{enumerate}[a)]
      \item Critical section.
      \item Atomic access.
    \end{enumerate}
\end{itemize}
\end{frame}

