\clearpage
\section{Submission}

The submission deadline for results of this lab will be announced in Aula Global.

Apply the following rules:

\begin{itemize}

  \item All submissions will be performed through Aula Global.

  \item You must deliver a zip file with the following files:
     \begin{itemize}
       \item \cppid{counter\_mutex.cpp}.
       \item \cppid{counter\_atomic.cpp}.
       \item \cppid{counter\_spin\_seq.cpp}.
       \item \cppid{counter\_spin\_seq\_opt.cpp}.
       \item \cppid{counter\_spin\_ra.cpp}.
       \item \cppid{counter\_spin\_ra\_opt.cpp}.
     \end{itemize}

  \item The content of the report must include:
  \begin{enumerate}
    \item Identifying data:
      \begin{itemize}
        \item Title.
        \item Name and NIA of the students.
        \item Reduced group where the students belong to.
        \item Lab group identifier.
      \end{itemize}

    \item Tables with results from evaluations.

    \item Brief conclusions.

  \end{enumerate}
  
\end{itemize}

The results obtained in the evaluations, will be included only in tabular form
with numeric information.  Clearly identify at the start of the section how
many times each experiment was executed.  Each table must have a row for each
program and a column for each number of threads (2, 4, 8, 16).

\textbad{IMPORTANT}: 
You must run a number of times which is enough to ensure an acceptable standard deviation.

You must include the following tables:

\begin{enumerate}
  \item Table with average execution times.
  \item Table with standard deviations.
\end{enumerate}
