\section{Advanced branch prediction techniques}

\begin{frame}[t]{Branch prediction}
\begin{itemize}
  \item High impact of \textgood{branches} on programs performance.

  \mode<presentation>{\vfill}
  \item To \textmark{reduce} impact:
    \begin{itemize}
      \item Loop unrolling.
      \item Branch prediction:
        \begin{itemize}
          \item Compile time.
          \item Each branch handled isolated.
        \end{itemize}
      \item Advanced branch prediction:
        \begin{itemize}
          \item Correlated predictors.
          \item Tournament predictors.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Branch dependence}

\begin{columns}[T]

\begin{column}{.4\textwidth}
\begin{itemize}
  \item If first and second branch are taken, third is \textbad{NOT-taken}.
\end{itemize}
\end{column}

\begin{column}{.3\textwidth}
\begin{block}{Example}
\lstinputlisting{int/m4-03-ilp/branch-pred-ex1.cpp}
\end{block}
\end{column}

\pause
\begin{column}{.3\textwidth}
\begin{block}{Example}
\lstinputlisting[language=generalasm2,basicstyle=\tiny]{int/m4-03-ilp/branch-pred-ex1.asm}
\end{block}
\end{column}

\end{columns}

\mode<presentation>{\vfill\pause}
\begin{itemize}
  \item Isolated prediction per individual branch does not give advantages.
    \begin{itemize}
      \item Need to predict depending on past history.
    \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[t]{Correlated prediction}
\begin{itemize}
  \item Maintains last branches \textmark{history} to select among several
        predictors.

  \mode<presentation>{\vfill\pause}
  \item A $(m,n)$ predictor:
    \begin{itemize}
      \item Uses the result of $m$ last branches to select among
            $2^m$ predictors.
      \item Each predictor has $n$ bits.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Predictor $(1,2)$:
    \begin{itemize}
      \item Result of last branch to select among 2 predictors.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Size of predictor}
\begin{itemize}
  \item A predictor $(m,n)$ has several entries for each branch address.

  \item Total size:
\[
S = 2^m \times n \times entries_{\text{address}}
\]

  \item Examples:
    \begin{itemize}
      \item $(0,2)$ with 4K entries $\rightarrow$ 8 Kb
      \item $(2,2)$ with 4K entries $\rightarrow$ 32 Kb
      \item $(2,2)$ with 1K entries $\rightarrow$ 8 Kb
   \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Miss rate}
\begin{itemize}
  \item \textbad{Activity}
    \begin{enumerate}
      \item \emph{Read} \textmark{Section 3.3} -- 
            Reducing Branch Costs With Advanced Branch Prediction.\\
            \textbad{Only} subsection \emph{Correlating Branch Predictors}. 
            (pg. 182--184 and Fig. 3.3).
        \begin{itemize}
          \item \bibhennessy
        \end{itemize}

      \mode<presentation>{\vfill}
      \item \textgood{Key aspects}:
        \begin{itemize}
          \item How do correlated predictors compare to simple predictors with the same size?
          \item How do correlated predictors to simple predictors with unlimited number of entries?
          \item What are the properties of \textmark{gshare} predictor?
        \end{itemize}
    \end{enumerate}
\end{itemize}
\end{frame}


\begin{frame}[t]{Tournament prediction}
\begin{itemize}
  \item \textgood{Combines two predictors}:
    \begin{itemize}
      \item \textmark{Global information} based predictor.
      \item \textmark{Local information} based predictor.
    \end{itemize}
  \item Uses a selector to choose between predictors.
    \begin{itemize}
      \item Change among two selectors uses a saturation counter (2 bits).
    \end{itemize}
  \item \textgood{Advantage}:
    \begin{itemize}
      \item Allows different behavior for integer and FP.
    \end{itemize}
  \item \textgood{SPEC}:
    \begin{itemize}
      \item \textmark{Integer benchmarks} $\rightarrow$ global predictor 40\%
      \item \textmark{FP benchmarks} $\rightarrow$ global predictor 15\%.
    \end{itemize}
  \item \textgood{Uses}: Alpha and AMD Opteron.
\end{itemize}
\end{frame}


\begin{frame}[t]{Intel Core i7}
\begin{itemize}
  \item \textbad{Activity}
    \begin{enumerate}
      \item \emph{Read} \textmark{Section 3.3} -- 
            Reducing Branch Costs With Advanced Branch Prediction.\\
            \textbad{Only} subsection 
            \emph{The Evolution of Intel Core i7 Branch Predictor}
            (pg. 190--191).
        \begin{itemize}
          \item \bibhennessy
        \end{itemize}
    \end{enumerate}
\end{itemize}
\end{frame}
