\section{Multiple issue techniques}

\begin{frame}[t]{CPI $<$ 1}
\begin{itemize}
  \item CPI $\geq$ 1 $\rightarrow$ Issue one instruction per cycle.

  \mode<presentation>{\vfill\pause}
  \item Multiple issue processors
        (\textmark{CPI} $<$ 1 $\rightarrow$ \textmark{IPC} $>$ 1):
    \begin{itemize}

      \mode<presentation>{\vfill}
      \item \textgood{Statically} scheduled superscalar processors.
        \begin{itemize}
          \item In-order execution.
          \item Variable number of instructions per cycle.
        \end{itemize}
      \mode<presentation>{\vfill}
      \item \textgood{Dynamically} scheduled superscalar \textgood{speculative} processors.
        \begin{itemize}
          \item Out-of-order execution.
          \item Variable number of instructions per cycle.
        \end{itemize}
      \mode<presentation>{\vfill}
      \item \textgood{VLIW} processors (\emph{Very Long Instruction Word}).
        \begin{itemize}
          \item Several instructions into a packet.
          \item Static scheduling.
          \item Hazard detection by the compiler.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Approaches to multiple issue}
\begin{itemize}
  \item Several approaches to multiple issue.
    \begin{itemize}
      \item \textmark{Static superscalar}: 
            Static scheduling with in-order execution.

      \mode<presentation>{\vfill\pause}
      \item \textmark{Dynamic superscalar}:
            Dynamic scheduling with out-of-order execution.

      \mode<presentation>{\vfill\pause}
      \item \textmark{Speculative superscalar}:
            Dynamic scheduling with out-of-order execution and speculation.

      \mode<presentation>{\vfill\pause}
      \item \textmark{LIW/VLIW/EPIC}:
            Primarily static scheduling with compiler support.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Superscalar processors}
\begin{enumerate}
  \item \textbad{Read} \textmark{Section 3.7} --
        Exploiting ILP using Multiple Issue and Static Scheduling
        (pg. 218--222).
    \begin{itemize}
      \item \bibhennessy
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Key aspects}:
    \begin{itemize}
      \item What are the key differences for each category of superscalar processor?
      \item Where are VLIW processors mostly used?
      \item What are the differences between VLIW and EPIC?
    \end{itemize}
\end{enumerate}
\end{frame}

