\section{Speculation}

\begin{frame}[t]{Branches and parallelism limits}
\begin{itemize}
  \item As parallelism increases, \textbad{control dependencies} become a harder problem.
    \begin{itemize}
      \item Branch prediction is not enough.
    \end{itemize}

  \mode<presentation>{\vfill}
  \item Next step is \textgood{speculation} on 
        \textmark{branch outcome} and \textgood{execution} 
        assuming that \textmark{speculation was right}.
    \begin{itemize}
      \item Instructions fetched, issued and executed.
      \item Need of a mechanism to handle wrong speculations.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Components}
\begin{itemize}
  \item \textgood{Ideas}:
    \begin{itemize}
      \item \textmark{Dynamic branch prediction}: 
            Selects instructions to be executed.
      \item \textmark{Speculation}: 
            Executes before control dependencies are resolved and may eventually undone.
      \item \textmark{Dynamic scheduling}.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item To achieve this, must \textgood{separate}:
    \begin{itemize}
      \item \textmark{passing instruction result} to another instruction using it.
      \item Instruction \textmark{finalization}.
    \end{itemize}

  \mode<presentation>{\vfill}
  \item \textmark{IMPORTANT}:
        Processor state (register file / memory) \textbad{not updated} until changes are confirmed.
\end{itemize}
\end{frame}

\begin{frame}[t]{Solution}
\begin{itemize}
  \item \textgood{Reorder Buffer} (\textmark{ROB}):
    \begin{itemize}
      \item When an instruction is finalized \textmark{ROB} is written.
      \item When execution is confirmed real target is written.
      \item Instructions read modified data from \textmark{ROB}.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textmark{ROB} entries:
    \begin{itemize}
      \item \textmark{Instruction type}: branch, store, register operation.
      \item \textmark{Target}: Register Id or memory address.
      \item \textmark{Value}: Instruction result value.
      \item \textmark{Ready}: Indication of instruction completion.
    \end{itemize}
\end{itemize}
\end{frame}
