\subsection{Benchmarks}

\begin{frame}[t]{Workload}
\begin{itemize}
  \item Computer \textmark{performance} depends on the
        evaluated \textmark{workload}.

  \mode<presentation>{\pause\vfill}
  \item Computers adapted to specific workloads:
    \begin{itemize}
      \item Web servers.
      \item Database servers.
      \item File servers.
      \item Personal computers.
      \item Multiprocessors.
      \item Multicomputers.
      \item \ldots
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Benchmarks}
\begin{itemize}
  \item Application or set of applications used to evaluate performance.

  \mode<presentation>{\pause\vfill}
  \item \textgood{Approaches}:
    \begin{itemize}
      \item \textmark{Kernels}: Small parts of real applications.
        \begin{itemize}
          \item \emph{Example}: FFT.
        \end{itemize}
      \item \textmark{Toy programs}: Short programs.
        \begin{itemize}
          \item \emph{Example}: Quicksort.
        \end{itemize}
      \item \textmark{Synthetic benchmarks}: 
            Invented to represent real applications.
        \begin{itemize}
          \item \emph{Example}: Dhrystone.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\pause\vfill}
  \item All are bad approaches:
    \begin{itemize}
      \item \alert{Architect and compiler might cheat!}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Benchmarks}
\begin{itemize}
  \item \textgood{Embedded}:
    \begin{itemize}
      \item Dhrystone (arguable relevance).
      \item EEMBC (kernels).
    \end{itemize}

  \mode<presentation>{\pause\vfill}
  \item \textgood{Desktop}:
    \begin{itemize}
      \item SPEC2017 (mix of integer and floating point programs).
    \end{itemize}

  \mode<presentation>{\pause\vfill}
  \item \textgood{Servers}:
    \begin{itemize}
      \item SPECWeb, SPECSFS, SPECjbb, SPECvirt\_Sc2010.
      \item TPC
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Example: SPEC2006}
\begin{itemize}
  \item \textmark{SPECrate/SPECspeed 2017 Integer}: Integer arithmetic.
    \begin{itemize}
      \item 20 programs: C (10), C++ (8), Fortran (2).
      \item Several application domains:
        \begin{itemize}
          \item Languages and compilers, 
                route planning,
                event simulation,
                video compression, 
                artificial intelligence, \ldots
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textmark{SPECrate 2017, SPECspeed 2017 Floating point}: Floating point
    \begin{itemize}
      \item 23 programs:
           Fortran (6),
           C (6),
           Fortran/C (5)
           C++ (2),
           C++/C (2),
           C++/C/Fortran (2).
      \item Several application domains:
        \begin{itemize}
          \item Physics, Chemistry, Biology, Algebra, image \emph{rendering},
                \ldots
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}
