\section{Resources}

\begin{frame}[t]{References}
\begin{itemize}
  \item \bulletenum{Basic reference}:
    \begin{itemize}
      \item \bulletmark{Computer Architecture: A quantitative approach, 6th Edition}. 
            Hennessy, JL and Patterson, DA. 
            Morgan Kaufmann, 2017.
    \end{itemize}
  \vspace{1em}
  \pause
  \item \bulletenum{Additional references}:
    \begin{itemize}
      \item \emph{Computer Organization and Design, MIPS Edition: The Hardware Software Interface}. 
            Patterson, DA; Hennessy, JL. 
            Morgan Kaufmann, 2020.
      \item \emph{The OpenMP Common Core: Making OpenMP Simple Again}.
            Mattson, TG; He, Y.; Koniges, A.E.
            MIT Press, 2019. 
      \item \emph{C++ Concurrency in Action. Practical Multithreading, 2nd Edition} 
            Williams, A.  
            Manning. 2018.
    \end{itemize}
\end{itemize}
\end{frame}

\ocwexclude{
\begin{frame}[t]{Other material}
\begin{itemize}
  \item Slides contents used in class will be published through \emph{Aula Global}.
  \vspace{1em}
  \item \alert{VERY IMPORTANT NOTICE}:
    \begin{itemize}
      \item Slides and other materials published through \emph{Aula Global} is just a \textbf{class outline}.
        \begin{itemize}
          \item \alert{They are not the course materials}
        \end{itemize}
      \item Only knowing the contents of those class outlines is insufficient for achieving the course goals.
        \begin{itemize}
          \item \alert{You are very likely to fail if you do not do more}.
        \end{itemize}
      \item It is highly recommended to use, study, and work with basic and additional references.
        \begin{itemize}
          \item \alert{e.g.: solve yourself recommended exercises from books}.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}
}

\ocwonly{
\begin{frame}{Course materials}
\begin{itemize}
  \item \bulletenum{Class materials}:
    \begin{itemize}
      \item Structured in 6 blocks.
      \item 15 lessons in total.
      \item At the end of each lesson.
        \begin{itemize}
          \item Additional recommended readings.
          \item Exercise recommendations.
        \end{itemize}
    \end{itemize}

  \vfill
  \item \bulletenum{Mandatory readings}:
    \begin{itemize}
      \item Complement to class materials.
    \end{itemize}

  \vfill
  \item \bulletenum{Exercises}:
    \begin{itemize}
      \item Collections for each thematic block.
    \end{itemize}

  \vfill
  \item \bulletenum{Labs}:
    \begin{itemize}
      \item Parallel and concurrent programming.
    \end{itemize}

  \vfill
  \item \bulletenum{Exams}.

\end{itemize}
\end{frame}
}

