\section{Conclusion}

\begin{frame}[t]{Summary}
\begin{itemize}
  \item SIMD approaches for increasing performance:
    \begin{itemize}
      \item Vector processors, SIMD extensions, GPGPU.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Vector architectures require specific hardware design.

  \mode<presentation>{\vfill\pause}
  \item SIMD extensions for multimedia popularized in x86.
    \begin{itemize}
      \item Challenges in autovectorization.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item GPGPU using a GPU as an accelerator.
    \begin{itemize}
      \item Programming model tied to hardware level.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{References}
\begin{itemize}
  \item \bibhennessy
    \begin{itemize}
      \item Section 4.1 -- \emph{Introduction}.
      \item Section 4.2 -- \emph{Vector Architecture}.
      \item Section 4.3 -- \emph{SIMD Instruction Set Extensions for Multimedia}.
      \item Section 4.4 -- \emph{Graphics Processing Units}.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{To be continued}
\begin{itemize}
  \item You will find more contents related to this course in:

  \mode<presentation>{\vfill\pause}
  \item \textgood{Concurrent and Parallel Programming}\\
        Elective course 4, 1C.
    \begin{itemize}
      \item Bachelor en Applied Mathematics and Computing.
      \item Bachelor in Computer Science and Engineering.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Computación de Altas Prestaciones}
    \begin{itemize}
      \item Máster en Ingeniería Informática.
      \item Máster en Ciencia y Tecnología Informática.
    \end{itemize}
\end{itemize}
\end{frame}
