\section{Other consistency models}

\begin{frame}[t]{Optimizations}
\begin{itemize}
  \item Models relaxing program execution order.
    \begin{itemize}
      \item Relax W $\rightarrow$ R $\Rightarrow$ 
            \textmark{Total Store Order}.
      \item Relax W $\rightarrow$ W and W $\rightarrow$ R $\Rightarrow$ 
            \textmark{Partial store order}.
      \item Relax all $\Rightarrow$ 
        \begin{itemize}
          \item \textmark{Weak ordering} (PowerPC).
          \item \textmark{Release consisntency} (RISC V).
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill}
  \item \textgood{Notation}:
    \begin{itemize}
      \item X $\rightarrow$ Y
        \begin{itemize}
          \item X must complete before Y is done.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Reorderings}
\mode<presentation>{\vfill}
\makebox[\textwidth][c]
{
\begin{tabular}{|l|c|l|}
\hline

Sequential consistency &
Opcionalmente &
R $\rightarrow$ R
R $\rightarrow$ W\\
& &
W $\rightarrow$ R
W $\rightarrow$ W
\\
\hline

Total store order &
IBM/370 &
R $\rightarrow$ R
R $\rightarrow$ W\\
&
DEC VAX &
W $\rightarrow$ W\\

& SPARC & \\
\hline

Partial store order &
SPARC &
R $\rightarrow$ R
R $\rightarrow$ W\\
\hline

Weak ordering &
Power PC &
\\
\hline

Release consistency &
MIPS, RISC V, ARM v8 &
\\
\hline


\end{tabular}
}

\end{frame}

\begin{frame}[t]{Reads bypass writes (W$\rightarrow$R)}
\begin{itemize}
  \item A \textgood{read} \textmark{may execute before} 
        a preceding \textgood{write}.

  \mode<presentation>{\vfill}
  \item Typical in systems with \textgood{write buffer}.
    \begin{itemize}
      \item Check consistency with buffer.
      \item Allow read buffer.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Other models}
\begin{itemize}
  \item R $\rightarrow$ W, W $\rightarrow$ R.
    \begin{itemize}
      \item Allow that \textgood{writes} 
            \textmark{may arrive into memory} out of program order.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item R $\rightarrow$ W, 
        W $\rightarrow$ R, 
        R $\rightarrow$ R, 
        W $\rightarrow$ W.
    \begin{itemize}
      \item Avoid only data and control dependencies within processor.
      \item \textgood{Alternatives}:
        \begin{itemize}
          \item Weak consistency.
          \item Release consistency.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Weak ordering}
\begin{itemize}
  \item Divides memory operations into \textgood{data operations} 
        and \textgood{synchronization operations}.

  \mode<presentation>{\vfill\pause}
  \item \textgood{Synchronization operations} act as a \textmark{barrier}.
    \begin{enumerate}
      \item All preceding data operations in program order to a synchronization 
            must complete before synchronization is executed.
      \item All subsequent data operations in program order to a synchronization 
            operation must wait until synchronization ins completed.
      \item Synchronization are performed in program order.
    \end{enumerate}

  \mode<presentation>{\vfill\pause}
  \item Hardware implementation of \textmark{barrier}.
    \begin{itemize}
      \item \textgood{Processor keeps a counter}:
        \begin{itemize}
          \item Data operation \textgood{issue} $\Rightarrow$ \textmark{increment}.
          \item Data operation \textgood{completed} $\Rightarrow$ \textmark{decrement}.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Release/acquire consistency}
\begin{itemize}
  \item \textmark{More relaxed} than weak consistency.
  \item  \textmark{Synchronization accesses} divided into:
    \begin{itemize}
      \item \textmark{Acquire}. 
      \item \textmark{Release}. 
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Semantics}:
    \begin{itemize}
      \item \textmark{Acquire}
        \begin{itemize}
          \item Must complete before all subsequent memory accesses.
        \end{itemize}
      \item \textmark{Release}
        \begin{itemize}
          \item Must complete all previous memory accesses.
          \item Subsequent memory accesses \textgood{MAY} initiate.
          \item Operations following a \textmark{release} and must wait, must be protected with an \textmark{acquire}.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}
