\section{Multi-cycle operations}

\begin{frame}[t]{Floating point operations}
\begin{itemize}
  \item One-cycle floating point operations?
    \begin{itemize}
      \item Having an extremely long clock cycle.
        \begin{itemize}
          \item Impact on global performance.
        \end{itemize}
      \item Very complex FPU control logic.
        \begin{itemize}
          \item Huge amount of resources for FP logic.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Alternative}: Floating point pipelining.
    \begin{itemize}
      \item Execution stage may be repeated several times.
      \item Multiple functional units in EX.
        \begin{itemize}
          \item \textmark{Example}: Integer unit, FP and integer multiplier, 
                FP adder, FP and integer divider.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Floating point pipeline}
\begin{itemize}
  \item EX stage now has a duration of more than 1 clock cycle.
\end{itemize}
\makebox[\textwidth][c]{
\input{en/m4-02-ilp/multicycle.tkz}
}

\end{frame}

\begin{frame}[t]{Latency and initiation interval}
\begin{itemize}
  \item \textgood{Latency}: Number of cycles between instruction producing
        the result and instruction using the result.
  \item \textgood{Initiation interval}: Number of cycles between issue of two instructions
        using the same functional units
\end{itemize}

\mode<presentation>{\vfill}
{
\begin{tabular}[c]{*{3}{|p{.3\textwidth}}|}
\hline
Operation & Latency & Initiation interval
\\
\hline
\hline

Integer ALU &
0  &
1
\\
\hline

Loads &
1 &
1
\\
\hline

FP addition &
3 &
1
\\
\hline

FP multiplication &
6 &
1
\\
\hline

FP division &
24 &
25
\\

\hline
\end{tabular}
}

\end{frame} 
