\section{Introduction to distributed shared memory}

\begin{frame}[t]{Snooping protocols and scalability}
\begin{itemize}
  \item \textbad{Problems} with snooping protocols.
    \begin{itemize}
      \item Requires communication with all caches:
        \begin{itemize}
          \item For each cache miss.
          \item For each shared data write.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item What is the \textgood{advantage} of snooping protocols?
    \begin{itemize}
      \item No centralized data structure.
        \begin{itemize}
          \item Low implementation cost.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item What is the \textbad{drawback} of snooping protocols?
    \begin{itemize}
      \item No centralized data structure.
        \begin{itemize}
          \item Communications limit scalability.
        \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{DSM basic model}
\makebox[\textwidth][c]{
\input{en/m5-02-dsm/dsm.tkz}
}
\mode<presentation>{\vfill\pause}
\begin{itemize}
  \item \textbad{Need to eliminate coherence traffic.}
\end{itemize}
\end{frame}

\begin{frame}[t]{Kinds of coherence protocols}
\begin{itemize}
  \item \textmark{Snooping} 
    \begin{itemize}
      \item Each cache keeps sharing status for each block it holds.
      \item Caches accessible through broadcasting medium (bus).
      \item All caches monitor if they have a copy of the block.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textmark{Directory based}:
    \begin{itemize}
      \item Sharing status kept in a directory.
        \item \textgood{SMP}: 
              Centralized directory in memory or last level cache (LLC).
        \item \textgood{DSM}: 
              To avoid bottleneck a distributed directory (more complex) is used.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Directory based protocol}
\begin{itemize}
  \item \textmark{Idea}: Keep \textgood{status} for each block in cache.
    \begin{itemize}
      \item Which caches have a block copy?
      \item Block state bits.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item Multi-cores with external shared cache.
    \begin{itemize}
      \item \textmark{Bit vector} with length equal to number of cores.
        \begin{itemize}
          \item Signals which private caches many have a copy of block.
          \item Only send invalidation to caches marked in bitmap.
        \end{itemize}
      \item This scheme works well within a single multi-core.
      \item \textmark{Example}: \textgood{Intel Core i7}.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Centralized directory and scalability}
\begin{itemize}
  \item A centralized directory avoids broadcasting but
    \begin{itemize}
      \item Becomes a bottleneck.
      \item Scalability problem with increasing number of processors.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Solution}: \textmark{Distributed directory}.
    \begin{itemize}
      \item Distribute directory with memory.
      \item Each directory has information about the associated local memory.
        \begin{itemize}
          \item Target directory is always known.
        \end{itemize}
      \item Different coherence requests go to different directories.
    \end{itemize}


\end{itemize}
\end{frame}

\begin{frame}{Distributed directory}
\makebox[\textwidth][c]{
\input{en/m5-02-dsm/dsm-dist.tkz}
}
\mode<presentation>{\vfill\pause}
\end{frame}
