\section{Cache coherence alternatives}

\begin{frame}[t]{Coherent multiprocessors}
\begin{itemize}
  \item A \textgood{coherent multiprocessor} offers:
    \begin{itemize}
      
      \mode<presentation>{\vfill\pause}
      \item Shared data \textmark{migration}.
        \begin{itemize}
          \item A datum may be moved to a local cache and be used transparently.
          \item Decreases remote data access latency and bandwidth demand to shared memory.
        \end{itemize}
      
      \mode<presentation>{\vfill\pause}
      \item Shared data \textmark{replication} simultaneously read.
        \begin{itemize}
          \item Performs data copy in local cache.
          \item Decreases access latency and read contention.
        \end{itemize}
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Critical properties for performance}:
    \begin{itemize}
      \item \textmark{Solution}: 
            Hardware protocol for keeping cache coherence.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Kinds of cache coherence protocols}
\begin{itemize}
  \item \textgood{Directory} based:
    \begin{itemize}
      \item Sharing state is kept in a directory.
      \item \textmark{SMP}: Centralized directory in memory or in LLC (\emph{Last-level cache}).
      \item \textmark{DSM}: To avoid bottlenecks a distributed directory is used (more complex).
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Snooping}:
    \begin{itemize}
      \item Each cache keeps the sharing state of each block that it stores.
      \item Caches accessible through a broadcasting device (bus).
      \item All caches monitor broadcasting device to determine if they have a copy of the block.
    \end{itemize}
\end{itemize}
\end{frame}
