\section{Barriers}

\begin{frame}[t]{Barrera}
\begin{itemize}
  \item A barrier allows to \textmark{synchronize} several processes in some point.
  
    \mode<presentation>{\vfill}
    \begin{itemize}
      \item Guarantees that no process passes the barrier until all have arrived.
      \mode<presentation>{\vfill}
      \item Used to synchronize phases in a program.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t]{Centralized barriers}
\begin{itemize}
  \item Centralized \textgood{counter} associated to the \textmark{barrier}.
    \begin{itemize}
      \item Counts number of processes that have arrived the barrier.
    \end{itemize}

  \mode<presentation>{\vfill\pause}
  \item \textgood{Barrier function}:
    \begin{itemize}
      \item Increments counter
      \item Waits the counter to reach the number of processes to be synchronized.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Simple barrier}
\begin{columns}

\column{.6\textwidth}
\begin{block}{Simple implementation}
\begin{lstlisting}
do_barrier(barrier, n) {
  lock(barrier.lock);
  if (barrier.counter == 0) {
    barrier.flag=0;
  }
  local_counter = barrier.counter++;
  unlock(barrier.lock);
  if (local_counter == NP) {
    barrier.counter=0;
    barrier.flag=1;
  }
  else {
    while (barrier.flag==0) {}
  }
}
\end{lstlisting}
\end{block}

\column{.4\textwidth}
\begin{itemize}
  \item Problem if barrier is reused in a loop.
\end{itemize}

\end{columns}
\end{frame}

\begin{frame}[t,fragile]{Barrier with way inversion}
\begin{columns}

\column{.6\textwidth}
\begin{block}{Simple implementation}
\begin{lstlisting}
do_barrier(barrier, n) {
  local_flag = !local_flag;
  lock(barrier.lock);
  local_counter = barrier.counter++;
  unlock(barrier.lock);
  if (local_counter == NP) {
    barrier.counter=0;
    barrier.flag=local_flag;
  }
  else {
    while (barrier.flag==local_flag) {}
  }
}
\end{lstlisting}
\end{block}

\column{.4\textwidth}

\end{columns}
\end{frame}

\begin{frame}[t]{Tree barriers}
\begin{itemize}
  \item A simple implementation of barriers is not \textbad{scalable}.
    \begin{itemize}
      \item Contention in access to shared variables.
    \end{itemize}
  \mode<presentation>{\vfill\pause}
  \item Tree structure for process arrival and release.
    \begin{itemize}
      \item Specially useful in distributed networks.
    \end{itemize}

\end{itemize}
\end{frame}
