\clearpage
\begin{acexercise}
\label{ex:m4-01:instr-03}
\end{acexercise}
%October 2013 exam.

A certain processor runs the following code segment:

\lstinputlisting[language=generalasm2]{int/ex/m4-01/instr3-code-question.asm}

Assume that the processor has 5 stages pipelined architecture (fetch,
decode, execute, memory and write-back) without forwarding. All stages
run in one cycle, except load and store operations which require two additional
cycles for memory access latency, and branching instructions which require
one additional execution cycle. Assume that both the banch target address
and the evaluation of the branch condition are known at the end of the 
decode stage.

\begin{enumerate}

  \item Identify RAW data hazards in the code.

  \item Show a timing diagram with execution stages for each instruction in
        one iteration.

  \item Determine how many cycles are required for executing one loop iteration
        when there is no branch prediction.

  \item Determine how many cycles are required for executing one loop iteration
        when a branch predictor (always predicting to taken) is used.

\end{enumerate}

\begin{acsolution}\end{acsolution}


\paragraph{Section 1}

\input{int/ex/m4-01/instr3-dep.tex}

\paragraph{Section 2}

Table~\ref{tab1:sol-oct-2013} shows the corresponding timing diagram.

\begin{table}[htb]
\input{int/ex/m4-01/instr3-time-01.tex}
\caption{Timing diagram for exercise~\ref{ex:m4-01:instr-03}.}
\label{tab1:sol-oct-2013}
\end{table}

\begin{itemize}

\item \asmlabel{I0}: Requires three memory cycles

\item \asmlabel{I1}: You can not start decoding until WB of \asmreg{t3} is done. Requires 3 cycles of memory.

\item \asmlabel{I2}: It can not be fetched until the stage is released by \asmlabel{I1}. You can not start to decode until the WB of \asmreg{t1}

\item \asmlabel{I4}: The decoding can not be started until \asmlabel{I2} does WB of \asmreg{t1}

\item \asmlabel{I5}: Fetch can not start until \asmlabel{I4} decoding starts. The memory cycle cannot be started  until the memory cycle of \asmlabel{I4} does not end.

\end{itemize}

\paragraph{Section 3}

In this case, instruction \asmlabel{I0} from next iteration cannot be
fetched until both the branch target address and the branch condition result
are known. This happens at the end of decode stage for \asmlabel{I5}. 
Consequently, \asmlabel{I0} is fetched in cycle 17 and one iteration
requires 16 cycles.

\paragraph{Section 4}

In this case, the prediction is performed in the decoding step, which is
when it is known that it is a branch instruction, so \asmlabel{I0}
starts in cycle 17, and one iteration requires 16 cycles.

\clearpage
