\begin{acexercise}\end{acexercise}
\label{ex:m4-01:instr-02}

Given the following code fragment:

\lstinputlisting[language=generalasm2]{int/ex/m4-01/instr2-question.asm}

A processor with a 5-stage pipeline is considered where hardware implements forwarding.
Instruction cache reads require one cycle. Data cache reads require two cycles and
writes require 3 cycles. For branch instructions both branch target address and
branch outcome are known at the end of execution stage and all branches are
always predicted to not taken.

Provide a time diagram (chronogram) for the first complete iteration
(including the fetch of the first instruction for next iteration).

\begin{acsolution}\end{acsolution}

In the given code the following RAW dependencies are identified:
\input{int/ex/m4-01/instr2-rawdep}

Solution is presented in table~\ref{ex:m4-01:instr-02:chrono:forward}.

Observations:

\begin{enumerate}

\item Stage \textmark{M} from instruction \asmlabel{I1}
      uses two cycles as it is a memory read.

\item A stall in instruction \asmlabel{I2} happens,
      as stage \textmark{M} is busy. 

\item Stage \textmark{M} from instruction \asmlabel{I2}
      uses two cycles as it is a memory read.

\item A stall in instruction \asmlabel{I3} happens,
      until the new value for \asmreg{t5} is available.
      The value is passed to stage \textmark{EX},
      through forwarding once it was obtained in stage \textmark{M} 
      from instruction \asmlabel{I2}.

\item A stall in instruction \asmlabel{I4} happens,
      as stage \textmark{ID} is busy. 

\item Stage \textmark{M} from instruction \asmlabel{I4}
      uses three cycles as it is a memory write.

\item A stall in instruction \asmlabel{I5} happens,
      as stage \textmark{M} is busy. 

\item A stall in instruction \asmlabel{I6} happens,
      as stage \textmark{EX} is busy. 

\item A stall in instruction \asmlabel{I7} happens,
      as stage \textmark{ID} is busy. 

\item A stall in instruction \asmlabel{I8} happens,
      as stage \textmark{IF} is busy. 

\item Instruction \asmlabel{I9} (next instruction) is fetched
      in cycle 14, as branch is predicted to not taken.

\item Branch is determined to be taken at the end of cycle 15.
      Consequently, instruction \asmlabel{I9} is retired from the pipeline
      and instruction \asmlabel{I1} is fetched.

\end{enumerate}

\begin{table}[htb]
\input{int/ex/m4-01/instr2-forward}
\caption{Time diagram for exercise~\ref{ex:m4-01:instr-02} wit forwarding.}
\label{ex:m4-01:instr-02:chrono:forward}
\end{table}
